Paper Title*\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2019-05-07 icdm/submission/includes/evaluation.tex
input shuffling

\subsection{Evaluating word embedings}
Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier. We cannot split our data set into train and test set. As the task that the network is learning is of no interest to us. Therefore, we need to verify that our embedding is of quality with other techniques. To define quality we first need to define a measure of similarity between two vectors. This requires knowledge of the Cosine distance, which is introduced in the following subsection.
\subsubsection{Cosine distance}
The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other. We get:
\begin{equation}
cos\_sim(v,w) = \frac{v \cdot w}{|v| |w|}
\end{equation}
The cosine of 0\textdegree ~is 1, it's 0 for two vectors that are orthogonal to each other and vectors that point in the opposite direction will have a cosine of their angle of -1. This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors, but by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors. This distance does not take into account any order of magnitude. Hence, for our tasks, two vectors will be considered equal if they are of different magnitude but point in the same direction.
Apart from measuring the quality of word embedding well, this technique has another advantage. By normalizing the vectors the calculation of the cosine angle becomes the dot product of the two vectors. Which can be computed very fast on modern GPU's.
Knowing that we have a measure to compute the similarity of two vectors let us introduce a way to rate the quality of our embeddings.

\subsubsection{Word similarity and wordsim353}
To measure the quality of our word embedding we will need a dataset to compare our results too. We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature. The data set consists of 353 pairs of words rated by humans on their similarity. The similarity score is in the range of 1 and 10, with 10 being the best score. An example for two of such pairs can be found in Figure \ref{fig:ws353_ex}. We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is a common procedure.

@@

\subsection{Configuration of the network}
The skip gram model has a lot of possible parameters, that can be tuned. We experimented with different models and finally decided for one that we tried to optimize. This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter. The explanation of the parameters will be structured as follows:
\texttt{Parameter} - Description and tuning - \textit{Value}
\begin{itemize}
\item \texttt{Negative Samples:} Here we have to find a trade-off between, setting the parameter too high which will result in increased accuracy but a longer computation time. For smaller data sets a higher number of negative samples is often needed. In their original paper Mikolov et al. \cite{mikolov2} recommend a value of 5-20. We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value. - \textit{10}
\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased. Mikolov et al. \cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5. - \textit{5}
\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}. We, therefore, used Gensim to find the best embedding possible. - \textit{100}
\item \texttt{Batch size}: As described in subsection \ref{ssec:b_SGM}, there is a tradeoff to find between quality and training time. We first used a batch size of 5000, but then decide after non conclusive results (see \ref{ssec:bs_lf}) that 2000 would be better - \textit{2000}
\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
\end{itemize}

\iffalse
\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
Embedding Size & Word Similarity on Gensim \\ \hline
50 & 0.65 \\ \hline
100 & 0.67 \\ \hline
200 & 0.65 \\ \hline
300 & 0.63 \\ \hline
\end{tabular}
\caption{Word similarity in relation to the size of the embedding}
\label{table:gensim_emb_size}
\end{table}
\fi
\subsection{Input Shuffling}
We used input shuffling as a technique to optimize the skip gram model. We will first describe input shuffling in a general way and then explain why we suppose that input shuffling could work well on the skip gram model. \\
Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as input at each epoch.
The idea behind this technique is to present our optimizer with different loss surfaces so that it's able to find the best optimum. Therefore, it's easier for the neural network to escape a local minimum. As for example if a network had converged to a local minimum after one epoch it could not escape it as all the parameters are the same. But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
\\
There are two reasons why we think that input shuffling is particularly well suited for the skip gram model. The first one has to do with the fact that when we read our words sequentially that words that only appear very early will not benefit from the context words being already updated from others. The second idea is that we used the special batch technique described in Section \ref{ssec:b_SGM}. When using this technique and not using shuffling we will always have words that appear next to each other in a batch and will, therefore, update the same words at the same time. We, therefore, lose some quality. But if we would use input shuffling instead, then the words in one batch would likely not be similar and therefore only taking the average of a small part of pairs with the same words will be less likely.

>>>>

\subsection{Evaluating word embedings}
Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier for example. We cannot split our data set into train and test set. As the task that the network is learning is of no interest to us. Therefore, to ensure the quality of word embeddings other techniques have to be used.To define quality we first need to define a measure of similarity between two vectors. This requires knowledge of the Cosine distance, which is introduced in the following subsection.
\subsubsection{Cosine distance}
The cosine similarity, this is not the cosine distance, of vectors $v$ and $w$ is the cosine of the angle between the two vectors It can be calculated by taking the dot product of $v$ and $w$ and dividing it by the magnitude of $v$ and $w$ multiplied with each other. We get:
\begin{equation}
cos\_sim(v,w) = \frac{v \cdot w}{|v| |w|}
\end{equation}
The cosine of 0\textdegree ~is 1, it's 0 for two vectors that are orthogonal to each other and vectors that point in the opposite direction will have a cosine of their angle of -1. This is not a good distance measure as -1 is smaller than 0, and therefore two vectors pointing away from each other would be closer than two orthogonal vectors, but by subtracting 1 from the cosine of the angle we can create a good distance measure between the two vectors. This distance does not take into account any order of magnitude. Hence, for our tasks, two vectors will be considered equal if they are of different magnitude but point in the same direction.
Apart from measuring the quality of word embedding well, this technique has another advantage. By normalizing the vectors the calculation of the cosine angle becomes the dot product of the two vectors. Which can be computed very fast on modern GPU's.
Knowing that we have a measure to compute the similarity of two vectors let us introduce a way to rate the quality of our embeddings.

\subsubsection{Word similarity and wordsim353}
To measure the quality of our word embedding we  use a dataset to compare our results too. We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature. The data set consists of 353 pairs of words rated by humans on their similarity. The similarity score is in the range of 1 and 10, with 10 being the best score. An example for two of such pairs can be found in Table \ref{fig:ws353_ex}. We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is the common procedure.

@@

\subsection{Configuration of the network}
The skip gram model has a lot of possible parameters, that can be tuned. We experimented with different models and finally decided for one that we tried to optimize. This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter. The explanation of the parameters will be structured as follows:
\texttt{Parameter} - Description and tuning - \textit{Value}
\begin{itemize}
\item \texttt{Negative Samples:} Here we have to find a trade-off between, a parameter that is too high (better accuracy) and low ( faster computation time). For smaller data sets a higher number of negative samples is often needed. In their original paper Mikolov et al. \cite{mikolov2} recommend a value of 5-20. We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value. - \textit{10}
\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased. Mikolov et al. \cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5. - \textit{5}
\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}. We, therefore, used Gensim to find the best embedding possible. - \textit{100}
\item \texttt{Batch size}: As described in Section \ref{ssec:b_SGM}, there is a trade off to find between quality and training time. We first used a batch size of 5000, but then decide after non conclusive results that 2000 would be better - \textit{2000}
\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
\end{itemize}

This conlcudes our overview of the possible parameters of our Skip-Gram Model, in the next section the reader will be given a description about the process of input shuffling.

\iffalse
\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
Embedding Size & Word Similarity on Gensim \\ \hline
50 & 0.65 \\ \hline
100 & 0.67 \\ \hline
200 & 0.65 \\ \hline
300 & 0.63 \\ \hline
\end{tabular}
\caption{Word similarity in relation to the size of the embedding}
\label{table:gensim_emb_size}
\end{table}
\fi
\subsection{Input Shuffling}
We used input shuffling as a technique to optimize the SGM.  We will first describe input shuffling in a general way and then explain why we supposed that input shuffling could work well on the SGM. \\
Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch.
The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in other to find the best optimum. Therefore, it's easier for the neural network to escape a local minimum. As for example if a network had converged to a local minimum after one epoch it probably cannot escape it. But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
\\
There are two reasons why we think that input shuffling is particularly well suited for the skip gram model. Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others. The second idea is that we used the special batch technique described in Section \ref{ssec:b_SGM}. When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples. But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller. In consequence the accuracy of the model should be optimized leading to a better convergence time. A definition of convergence time is given in the next section.
